Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-10-25 GWLDCII.tex
last 2 paras of intro

This paper continues this body of work by establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. This apparent contradiction led to the work in \cite{non-orientable}, which showed that the actual geometric space defined by the Wilson loop diagrams is a non-orientable $k$-dimensional bundle over this $3k$-dimensional manifold in $\Gr(k,n)$.

SYM $N=4$ is a finite theory \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}. That is, the scattering amplitudes for particle interactions are finite. Since the (sum of) integrals associated to these scattering amplitudes can be associated to (the union of) convex polytopes, as the  noted above, one considers the sum of integrals as defining a volume to said union of polytopes. However, each individual integral has singularities occuring on the boundaries of the corresponding convex polytope. If the singularities cancel out in the overall sum, then the overall volume is finite. This cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general. In this paper, we give a goemetric characterization of the factors appearing in the denominator of these integrals.  The set of irreducible factors in the denominator, as defined in the physics literature \cite{HeslopStewart}, are exactly the irreducible factors of the minors defined by the Grassmann necklace in a physically significant representation of the positroid cell assoicated to a diagram. In other words, the singluarities arising from tree level particle interactions in SYM $N=4$ theory do not just come from the Feynman rules describing the theory. They can also be seen directly from the associated geometry, given the physically correct representation.  We hope these results will pave the way for a general study of these poles in future.

>>>>

This paper continues this body of work by establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. This apparent contradiction is resolved by the non-orientability result of \cite{non-orientable}.

As described above, each integrand associated to a Wilson loop diagram gives the volume of a convex polytope.  These individual integrals have singularities which come from the factors in the denomiator of the integrands. However SYM $N=4$ is a finite theory \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}, so the singularities must cancel out in the sum over all diagrams. This cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$-dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto \cite{GWLDCI}.

In this paper, we give a geometric characterization of the factors appearing in the denominator of these Feynman integrals in terms of the Grassmann necklace of the associated positroid. Specifically, we show that the set of irreducible factors in the denominator, as defined by the physics \cite{HeslopStewart}, are exactly the irreducible factors of the minors defined by the Grassmann necklace evaluated on a physically significant realization of the associated positroid cell. In other words, the singularities arising from tree level particle interactions in SYM $N=4$ theory do not just come from the Feynman rules describing the theory: they can also be seen directly from the associated geometry, given the physically correct representation.  We hope these results will pave the way for a general study of these poles in future.

%SYM $N=4$ is a finite theory \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}. That is, the scattering amplitudes for particle interactions are finite. Since the (sum of) integrals associated to these scattering amplitudes can be associated to (the union of) convex polytopes, as the  noted above, one considers the sum of integrals as defining a volume to said union of polytopes. However, each individual integral has singularities occuring on the boundaries of the corresponding convex polytope. If the singularities cancel out in the overall sum, then the overall volume is finite. This cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general. In this paper, we give a goemetric characterization of the factors appearing in the denominator of these integrals.  The set of irreducible factors in the denominator, as defined in the physics literature \cite{HeslopStewart}, are exactly the irreducible factors of the minors defined by the Grassmann necklace in a physically significant representation of the positroid cell assoicated to a diagram. In other words, the singluarities arising from tree level particle interactions in SYM $N=4$ theory do not just come from the Feynman rules describing the theory. They can also be seen directly from the associated geometry, given the physically correct representation.  We hope these results will pave the way for a general study of these poles in future.
